---
title: "R Notebook"
output: 
  html_notebook: 
    toc: yes
runtime: shiny 
---

# Set up

As a first step in the analysis, we need to load a count matrix.

This matrix will be used by DESeq2 to perform its differential analysis, to find genes that are differentially expressed between conditions.

On the IFB RStudio server, data is available here: "/shared/data/projects/fair_training2020/Result/counts_matrix.txt"

Copy it first to your home folder, along with the R script

```{r}
counts <- read.table("./counts_matrix.txt", sep="\t", row.names="GeneId", header=TRUE)
```

# Data exploration

A first look at the count matrix. We have 6 samples.

```{r, echo = FALSE}
numericInput("n",   "How many lines ?", 5)
renderTable({   head(counts, input$n) })
```


# Data preparation

Then we need to provide the experimental design so that DESeq2 knows which sample groups to compare.

```{r}
sampleInfo <- data.frame(condition = c("iron", "iron", "iron","noIron", "noIron", "noIron"),
                     row.names = names(counts))
sampleInfo
```

# DESeq2 analysis

Next step: load the data into an object useable by DESeq2.

```{r}
library("DESeq2")
dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = sampleInfo,
                              design = ~ condition)
dds
```

Differential expression analysis

```{r}
# Perform the analysis by providing the data to DESeq
dds <- DESeq(dds)
# Extract the results
res <- results(dds)
# Display the results
# Re-order by adjusted p-value
resOrdered <- res[order(res$padj),]
resOrdered
```

# Results

Display a summary of the results

```{r}
summary(resOrdered)
```

Display a MA plot of the dataset. Statistically significant genes are displayed in red.

```{r}
plotMA(res)
```

# Session info

Keep versions of the packages used for the analysis

```{r}
sessionInfo()
```

